---
title: "320 Homework 2"
author: "Max Sleek"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

**Note: Throughout this document, I included the head() function, because if I didn't, R would knit and produce all 336,776 rows of flights, causing my knitted document to be 16,000 pages long.**

# Problems:
  
```{r}
# Section 5.2.4 exercises 1-3; 
# Section 5.3.1 exercises 2-4;
# Section 5.4.1 exercises 2-4.
```

# Packages and Stuff
```{r}
library(ggplot2)
library(nycflights13)
library(tidyverse)
flights <- data.frame(flights)
as_tibble(flights)
```

# Section 5.2.4

**1. Find all flights that: **

```{r}
# Had an arrival delay of two or more hours
head(filter(flights, arr_delay >= 120))
# Flew to Houston (IAH or HOU)
head(filter(flights, dest == "IAH" | dest == "HOU"))
# Were operated by United, American, or Delta
head(filter(flights, carrier == "UA" | carrier == "AA" | carrier == "DL"))
# Departed in summer (July, August, and September)
head(filter(flights, month %in% c(7, 8, 9)))
# Arrived more than two hours late, but didn’t leave late
head(filter(flights, arr_delay >120 & dep_delay <=0))
# Were delayed by at least an hour, but made up over 30 minutes in flight
head(filter(flights, dep_delay >=60, arr_delay <30))
# Departed between midnight and 6am (inclusive)
head(filter(flights, dep_time >=0, dep_time <=600))
```

**2. Another useful dplyr filtering helper is between(). What does it do? Can you use it to simplify the code needed to answer the previous challenges?**

The between function can be used within the filter function to return a subset of a data frame within certain boundaries. Yes, we can use it to simplify the previous code. For example:

```{r}
head(filter(flights, between(month, 7, 9)))
```

**3. How many flights have a missing dep_time? What other variables are missing? What might these rows represent?**

```{r}
head(filter(flights, is.na(dep_time)))
# 8255 values missing dep_time
Missing_Values <- is.na.data.frame(flights)
# By saving the is.na function to a matrix, I was able to sort it and detect which variables had missing values. I'm sure there is an easier way, but I found that the other variables were: dep_delay, arr_time, arr_delay, and air_time. Since these are all time-related values, these rows might represent flights that never took off, or crashed during flight. 
```

**4. Why is NA ^ 0 not missing? Why is NA | TRUE not missing? Why is FALSE & NA not missing? Can you figure out the general rule? (NA * 0 is a tricky counterexample!)**

(I accidentally did #4, but why not leave it in for practice). NA ^ 0 is not missing because any number to the power of 0 is 1. If we assign x to NA ^ 0, x will be stored as 1. NA | TRUE is not missing because x is stored as TRUE, not NA. Similarly, FALSE & NA is not missing because x is stored as FALSE. see below.

```{r}
x <- NA ^ 0
x
x <- NA | TRUE
x
x <- FALSE & NA
x
```


# Section 5.3.1

**2. Sort flights to find the most delayed flights. Find the flights that left earliest.**

```{r}
# Most delayed flights
head(arrange(flights, desc(dep_delay)))
# Flights that left earliest
head(arrange(flights, dep_delay))
```

**3. Sort flights to find the fastest (highest speed) flights.**

```{r}
head(arrange(flights, desc(distance/air_time)))
```

**4. Which flights traveled the farthest? Which traveled the shortest?**

```{r}
# Farthest
head(arrange(flights, desc(distance)))
# Shortest
head(arrange(flights, distance))
```

# Section 5.4.1

**2. What happens if you include the name of a variable multiple times in a select() call?**

Nothing happens, the variable only appears once. See Below.

```{r}
head(dplyr::select(flights, year, dep_delay, year))
```

**3. What does the any_of() function do? Why might it be helpful in conjunction with this vector? vars <- c("year", "month", "day", "dep_delay", "arr_delay")**

The any_of() function operates within a select function. If you store the desired variable names in a vector, then place the name of the vector in any_of(), those variables will be selected from the data frame. This is useful when there is many variables/columns. See Below.

```{r}
vars <- c("year", "month", "day", "dep_delay", "arr_delay")
head(dplyr::select(flights, any_of(vars)))
```

**4. Does the result of running the following code surprise you? How do the select helpers deal with case by default? How can you change that default?**

I'm not surprised that this argument exists, though it is a bit surprising that the case of "TIME" doesn't matter. The select helpers are defaulted to ignore case. This can be fixed with the code below.

```{r}
head(select(flights, contains("TIME")))
# Make select helpers case sensitive
head(select(flights, contains("time", ignore.case = FALSE)))
```

