---
title: "320 Homework 3"
author: "Max Sleek"
date: "`r Sys.Date()`"
output: html_document
---

# Problems:
  
```{r}
# Section 7.3.4 Questions 1, 3; 
# Section 7.4.1 Questions 1;
# Section 7.5.1.1 Question 2;
# Section 7.5.2.1 Question 1;
# Section 7.5.3.1 Question 2,3,5;
```

# Packages and Stuff
```{r}
library(ggplot2)
library(tidyverse)
diamonds <- data.frame(diamonds)
diamonds <- as_tibble(diamonds)
```

# Section 7.3.4

**1. Explore the distribution of each of the x, y, and z variables in diamonds. What do you learn? Think about a diamond and how you might decide which dimension is the length, width, and depth.**

Each of the distrobutions has a semi-normal shape with a skew towards the right. The scales of the historgrams are distorted because of the outliers to the right. This is likely because of the few exorbatant diamonds purchased by the wealthy. For dimensions, x and y may represent length and width becuase they are similar, and y may represent depth because it is a lower value than x and y.

```{r}
ggplot(data = diamonds) +
  geom_histogram(mapping = aes(x = x), binwidth = 0.1)

ggplot(data = diamonds) +
  geom_histogram(mapping = aes(x = y), binwidth = 0.1)

ggplot(data = diamonds) +
  geom_histogram(mapping = aes(x = z), binwidth = 0.1)
```

**3. How many diamonds are 0.99 carat? How many are 1 carat? What do you think is the cause of the difference?**

There are 23 0.99 carat diamonds and 1,558 1 carat diamonds. I think this is the case because people are much more likely to buy the whole number because it sounds better to themselves and their significant other.

```{r}
diamonds %>%
  count(carat == 0.99)

diamonds %>%
  count(carat == 1.0)
```

# Section 7.4.1

**1. What happens to missing values in a histogram? What happens to missing values in a bar chart? Why is there a difference?**

In a histogram, ggplot will remove the missing values and provide a warning when the code is ran. In a bar chart, ggplot will separate the missing values into their own "NA" column at the end of the chart. The difference is present becuase values in a histogram MUST be numeric. They cannot be plotted if they are missing. Since bar charts use categorical data, a column can be created where the missing values are treated as their own categorical variable.

# Section 7.5.1.1

**2. What variable in the diamonds dataset is most important for predicting the price of a diamond? How is that variable correlated with cut? Why does the combination of those two relationships lead to lower quality diamonds being more expensive?**

I struggled with this question. The carat variable is the most important predictor for price. Using the frequncy polygon plot, we can find that cut and carat have a weak, slightly negative correlation. Using the final boxplot, we can see that there is a lot of variability within each cut type, but the weak negative correlation is still maintained. I can figure that this relationship exists because smaller diamonds probably require a better quality cut, while larger ones can be cut with less quality.

```{r}
# Relationship between carat and price
ggplot(data = diamonds, aes(x = carat, y = price)) +
  geom_point(aes(color = cut))

ggplot(data = diamonds, aes(x = carat, y = price)) +
  geom_boxplot(aes(group = cut_width(carat, 0.1)))

# Relationship between carat and cut
ggplot(data = diamonds, mapping = aes(x = carat, y = ..density..)) + 
  geom_freqpoly(mapping = aes(color = cut))

ggplot(data = diamonds, aes(x = cut, y = carat)) +
  geom_boxplot()
```

# Section 7.5.2.1

**1. How could you rescale the count dataset above to more clearly show the distribution of cut within colour, or colour within cut?**

```{r}
# Cut within Color
diamonds %>% 
  count(color, cut) %>% 
  group_by(color) %>%
  mutate(ratio = n / sum(n)) %>%
  ggplot(aes(x = color, y = cut)) +
  geom_tile(aes(fill = ratio))

# Color within Cut
diamonds %>% 
  count(color, cut) %>% 
  group_by(cut) %>%
  mutate(ratio = n / sum(n)) %>%
  ggplot(aes(x = color, y = cut)) +
  geom_tile(aes(fill = ratio))
```

# Section 7.5.3.1

**2. Visualise the distribution of carat, partitioned by price.**

```{r}
ggplot(data = diamonds, mapping = aes(x = price, y = carat)) + 
  geom_boxplot(mapping = aes(group = cut_number(price, 10)))

# You can also flip the coordinates to help with the aesthetics of the graph, I learned this off of the internet
ggplot(data = diamonds, mapping = aes(x = price, y = carat)) + 
  geom_boxplot(mapping = aes(group = cut_number(price, 10))) +
  coord_flip()
```

**3. How does the price distribution of very large diamonds compare to small diamonds? Is it as you expect, or does it surprise you?**

The minimum prices for larger diamonds is higher than smaller diamonds, which I woould expect. The median is also higher, with the highest values in the entire dataset serving as the outliers. I would expect these results, as a larger diamond should cost more than a smaller one.

**5. Two dimensional plots reveal outliers that are not visible in one dimensional plots. For example, some points in the plot below have an unusual combination of x and y values, which makes the points outliers even though their x and y values appear normal when examined separately. Why is a scatterplot a better display than a binned plot for this case?**

A scatterplot is better than a binned plot in this case because we can see where the individual points lay comapred to the cluster. While binned plots ARE two dimensional, they often lack the ability to show singular values and their specific location. They often blend in with the rest of the dataset. 

```{r}
# Code referenced in the question
ggplot(data = diamonds) +
  geom_point(mapping = aes(x = x, y = y)) +
  coord_cartesian(xlim = c(4, 11), ylim = c(4, 11))

# Example binned plot
ggplot(data = diamonds) +
  geom_hex(mapping = aes(x = x, y = y)) +
  coord_cartesian(xlim = c(4, 11), ylim = c(4, 11))
```







