---
title: "320 HW 1"
author: "Max Sleek"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

# Problems:
  
```{r}
  # Section 3.2.4 Exercises 1-5; 
  # Section 3.3.1 Exercises 1-2, 5; 
  # Section 3.6.1 Exercises 1-5. 
```

# Packages
```{r}
library(ggplot2)
```

# Section 3.2.4

#1: Run ggplot(data = mpg). What do you see?

```{r}
ggplot(data=mpg)
  # The output is a blank graph. More arguments are needed.
```

#2: How many rows are in mpg? How many columns?
 
```{r}
nrow(mpg)
ncol(mpg)
str(mpg)
  # There are 234 rows and 11 columns. 
```

#3: What does the drv variable describe? Read the help for ?mpg to find out.

```{r}
?mpg 
  # By using the help function for mpg, we can see that drv describes that type of drive train in the vehicle.
```

#4: Make a scatterplot of hwy vs cyl.

```{r}
ggplot(data=mpg) +
  geom_point(mapping=aes(x=cyl, y=hwy))
```

#5: What happens if you make a scatterplot of class vs drv? Why is the plot not useful?

```{r}
ggplot(data=mpg) +
  geom_point(mapping=aes(x=drv, y=class))
  # Since drv and class are categorical variables, not numeric, a scatterplot is not useful. We cannot see how many are in each category.
```

# Section 3.3.1

#1: What’s gone wrong with this code? Why are the points not blue?
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
  # The color argument needs to be outside of the aes function. The color argument is an argument of the geom_point function. The proper code would be:
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```

#2: Which variables in mpg are categorical? Which variables are continuous? (Hint: type ?mpg to read the documentation for the dataset). How can you see this information when you run mpg?
```{r}
str(mpg)
  #By running the str function, we can see the type of variable and a few examples of observations. Manufacturer, model, trans, drv, fl, and class are categorical. Displ, year, cyl, cty, and hwy are continuous. 
```

#5: What does the stroke aesthetic do? What shapes does it work with? (Hint: use ?geom_point)

```{r}
# The stroke aesthetic works on shapes that have a border. It modifies the width of that border.
```

# Section 3.6.1

#1: What geom would you use to draw a line chart? A boxplot? A histogram? An area chart?
  
```{r}
  # Line chart: geom_line
  # Boxplot: geom_boxplot
  # Histogram: geom_histogram
  # Area chart: geom_area
```

#2: Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = TRUE)
  # I think that a scatterplot will be created, though I'm not sure what geom_smooth does.
  # geom_smooth created lines to help with visuals.
```

#3: What does show.legend = FALSE do? What happens if you remove it? Why do you think I used it earlier in the chapter?
  
```{r}
# show.legend removes the legend from the plot. without the show.legend function, R assumes the function to be true ans hows the legend. show.legend can be used when the data is purely numerical and does not to be separated. A legend would not be necessary.
```

#4: What does the se argument to geom_smooth() do?
  
```{r}
# It shows (or doesn't show) the confidence interval around the smooth function's output.
```

#5: Will these two graphs look different? Why/why not?

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))

# They will look the same. The arguments used in the different pieces of code can be moved around while still maintaining the same output.
```

























