---
title: "Analysis 2: Connecting Country Level Data"
author: "Max Sleek"
date:  "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

# Instructions

**Overview:** For each question, show your R code that you used to answer each question in the provided chunks. When a written response is required, be sure to answer the entire question in complete sentences outside the code chunks. When figures are required, be sure to follow all requirements to receive full credit. Point values are assigned for every part of this analysis.

**Helpful:** Make sure you knit the document as you go through the assignment. Check all your results in the created HTML file.

**Submission:** Submit via an electronic document on Sakai. Must be submitted as an HTML file generated in RStudio. 

```{r setup, include=F}
knitr::opts_chunk$set(echo = TRUE)
options(scipen=999)
library(tidyverse) 
library(rvest)
library(curl) #Make sure you install this package if you don't have it.
```

# Introduction

Many times in data science, your data will be split between many different sources, some of which may be online. In this analysis assignment, we will webscrape country level data from multiple websites, clean the data individually, and merge the data. The website [Worldometers](https://www.worldometers.info/) contains very interesting country level data that when connected may allow us to learn interesting things about the wonderful world in which we exist:

Follow the steps to accomplish specific tasks, and do not modify code on lines with the comment **#DO NOT CHANGE**. Add code in R code chunks wherever you see COMPLETE. After completing each code chunk, change `eval=FALSE` to `eval=TRUE` and **knit** the document.

# Assignment

## Part 1: GDP by Country

Information at [Worldometer GDP](https://www.worldometers.info/gdp/gdp-by-country/) contains GDP data from 2017 published by the world bank. GDP is the monetary value of goods and services produced within a country over a period of time. On this website, GDP is presented in dollars.
 
### Q1 *(3 Points)*

Webscrape the data from https://www.worldometers.info/gdp/gdp-by-country/ into a data frame in R called `GDP`. If done correctly, you should have a new object in R called `GDP` which is a data frame with 189 observations and 8 variables.

```{r,eval=TRUE}
URL.GDP="https://www.worldometers.info/gdp/gdp-by-country/" #DO NOT CHANGE

GDP = read_html(URL.GDP) %>%
  html_table() %>%
  .[[1]]

head(GDP) #DO NOT CHANGE
```

### Q2 *(2 Points)*

Remove the first and fourth variables from `GDP` and create a new data frame named `GDP2` based on this change.

```{r,eval=TRUE}
GDP2 = GDP %>% subset(select = -c(1,4))

head(GDP2) #DO NOT CHANGE
```

### Q3 *(3 Points)*

Create a new data frame named `GDP3` based off `GDP2` where the variables `GDP (nominal, 2017)`,`GDP  growth`, `Population (2017)`, `GDP  per capita`, and `Share of World GDP` are renamed to `GDP`, `Growth`, `Population`, `PerCapita`, and `Share`, respectively. Be careful!! In the original variable names, there are two spaces between "GDP" and "growth" in `GDP growth` and two spaces between "GDP" and "per capita" in GDP  per capita`.

```{r,eval=TRUE}
GDP3 = GDP2 %>% rename("GDP" = `GDP (nominal, 2017)`, "Growth" = `GDP  growth`, "Population" = `Population (2017)`, "PerCapita" = `GDP  per capita`, "Share" = `Share of World GDP`)

names(GDP3) #DO NOT CHANGE
```


### Q4 *(3 Points)*

Next, we must clean the data so there are no dollar signs or percent signs in the data using `str_replace()`. The dollar sign is a special character and must be referenced as `\\$`. Create a new data frame named `GDP4` where the dollar signs and percent signs are removed from all necessary variables.

```{r, eval=TRUE}
GDP4 = GDP3
GDP4$GDP = str_replace(GDP4$GDP, "\\$", "")
GDP4$PerCapita = str_replace(GDP4$PerCapita, "\\$", "")
GDP4$Growth = str_replace(GDP4$Growth, "%", "")
GDP4$Share = str_replace(GDP4$Share, "%", "")

str(GDP4) #DO NOT CHANGE
```

### Q5 *(3 Points)*

Next, create a new data frame named `GDP5` where all commas are removed from potentially numeric variables using `str_replace_all()`.

```{r, eval=TRUE}
GDP5 = GDP4
GDP5$GDP = str_replace_all(GDP5$GDP, ",", "")
GDP5$Population = str_replace_all(GDP5$Population, ",", "")
GDP5$PerCapita = str_replace_all(GDP5$PerCapita, ",", "")

str(GDP5) #DO NOT CHANGE
```

### Q6 *(2 Points)*

Create a new data frame called `GDP6` where all the variables except `Country` are changed to numeric variables.

```{r, eval=TRUE}
GDP6 = (GDP5)
class(GDP6$GDP) = "numeric"
class(GDP6$Growth) = "numeric"
class(GDP6$Population) = "numeric"
class(GDP6$PerCapita) = "numeric"
class(GDP6$Share) = "numeric"

str(GDP6) #DO NOT CHANGE
```


### Q7 *(2 Points)*

Rewrite over the original `GDP` variable with a new variable called `GDP` that is in **trillions of dollars** rather than in actual dollars. Rewrite over the original `Population` variable with a new variable of the same name that is in **millions of people** rather than in actual people. You are scaling the original variables to change the units without changing the variable names. Save your changes in a new data frame called `GDP7`.

```{r, eval=TRUE}
GDP7 = GDP6
GDP7$GDP = GDP7$GDP/1000000000000
GDP7$Population = GDP7$Population/1000000

str(GDP7)  #DO NOT CHANGE
```

## Part 2: More Relevant Data for Countries

Check out the Wikipedia page (https://en.wikipedia.org/wiki/Education_Index) which contains the education index for all countries from 1990 to 2019.

Check out the Johns Hopkins page (https://coronavirus.jhu.edu/data/mortality) which contains covid data for all countries.

### Q1 *(4 Points)*

Webscrape the data from https://en.wikipedia.org/wiki/Education_Index into a data frame in R called `EDU` and only keep the country's name and educational index from 2019.  

```{r,eval=TRUE}
URL.EDU="https://en.wikipedia.org/wiki/Education_Index" #DO NOT CHANGE

EDU = read_html(URL.EDU) %>%
  html_table() %>%
  .[[1]] %>%
  subset(select = c(`Country`, `2019`))

head(EDU) #DO NOT CHANGE
```


### Q2 *(4 Points)*

Webscrape the data from https://coronavirus.jhu.edu/data/mortality into a data frame in R called `COVID` and make the variables named `Confirmed`, `Deaths`, and `Case-Fatality` numeric by first removing the commas and percent signs and then by using the `as.numeric()` function.

Each time you run this code chunk, you may have to run this multiple times to get it to work. You may get some error related to "open.connection".  You can also try knitting the document every time you make any change to this code chunk to see the output in the HTML file.

```{r,eval=FALSE}

URL.COVID="https://coronavirus.jhu.edu/data/mortality" #DO NOT CHANGE

COVID = read_html(URL.COVID) %>%
  html_nodes("MortalityRates_container__2S9lV")

str(COVID) #DO NOT CHANGE
```

## Part 3: Merging the Datasets and Follow-up 

Now, we are going to merge the datasets for maximum gains. Make sure you carefully read the instructions for each question. Be very careful in this part of the assignment.

### Q1 *(2 Points)*

The dataset named `GDP7` is our primary dataset. Create a new dataset called `GDP7.EDU` that brings the education data from `EDU` into the dataset `GDP7`. Use a left join only.

```{r,eval=TRUE}
GDP7.EDU = left_join(GDP7, EDU)

head(GDP7.EDU) #DO NOT CHANGE
```


### Q2 *(3 Points)*

How many countries in `GDP7.EDU` have missing values for Education Index? Show code that can be used to answer this question and place your answer in the appropriate spaces below.

Code *(1 Point)*:
```{r,eval=TRUE}
length(which(is.na(GDP7.EDU$`2019`)))
```

Answer *(2 Point)*: There are 24 countries in the `GDP7.EDU` dataset that have missing values for Education Index.


### Q3 *(2 Points)*

Suppose we want to see if countries with higher GDP are less affected by the coronavirus. Merge the data from `GDP7` into the the data frame named `COVID` to create a new dataset named `GDP7.COVID`. The dataset `COVID` is now our primary dataset. Use a right join only.

```{r,eval=FALSE}
COMPLETE

head(GDP7.COVID) #DO NOT CHANGE
```