---
title: "320 Lab 7"
author: "Max Sleek"
date: "`r Sys.Date()`"
output: html_document
---

The inference that we have seen within the Modeling 1 lecture has all been catered to means.  However, recall from 155 that we also have analogous procedures for inference for proportions.  This lab will elucidate exactly these procedures.  

To this end, recall the Central Limit Theorem for one proportion.  That is, for indepedent observations of a sufficient sample size the sample proportion has an unbiased and normal sampling distribution.  More carefully, $\hat{P} \sim N(P, \sqrt{\frac{P(1-P)}{n}})$.  Often times in practice we do not have the true P, so we use the substitute the estimator $\hat{P}$ to estimate this standard error.  That said, when the null hypothesis confers upon us an assumption about the true $P$, we are at liberty to leverage this information.  

# 

Write a function, similar to `ttest` in the Modeling 1 lecture, called `ztest`.  I have included for you the name as well as the necessary arguments.  This function should take the number of observations adhering to the proportion criteria, $x$, as well as the sample size and compute a sample proportion, $\hat{p}$.  Then it should use this point estimate to create a test statistic and corresponding two sided p-value.  The function should then conclude according to this p-value (2pt).  

Remember that you will have to delete the `eval = FALSE` component of the r header to execute these code chunks.  


```{r, eval = TRUE}
ztest = function(x, n, h0, alpha){
  phat = x/n
  z = (phat - h0)/(sqrt((h0*(1-h0))/n))
  pval = 2 *(pnorm(-abs(z)))
  conclusion = if (pval<alpha){
    "Reject Null"
  }else{
    "Cannot Reject Null"
  }
  return(list(conclusion, pval, z))
}
```

#

I have a belief that roughly 15\% of the people in my graduate department have blue eyes.  In reality, out of a department with 105 students, 18 have blue eyes.  Formulate this scenario as a hypothesis test and submit it to the `ztest` function at the $0.01$ significance level (1pt) . 

```{r}
ztest(18, 105, 0.15, 0.01)
```

#

You should have gotten a conclusion of failing to reject the null.  Write a *loop* that will compute how many *more* student would need to have blue eyes before I had sufficient evidence to reject the null.  You can use hard coding to check your answer, but the actual procedure itself should be soft coded as a loop (2pt).  


```{r, eval = TRUE}
#initialize variables
x = 18
alpha = 0.01
pval = 0.5385
while(pval>alpha) {
  x = x + 1
  pval = ztest(x, 105, 0.15, alpha)[2]
}

print(pval)
x-18
# 8 Students
```

#

R has built-in functions to accomplish this task as well.  When the sample size is sufficiently large -- $P, 1-P \geq 10$ -- the `prop.test` function can be used analogously to our function above.  It structure is as follows: `prop.test(x, n, p = NULL, alternative = "two.sided", correct = TRUE)`.  Note, the correction referenced is the Yates correction and should be set to FALSE to get a mathematically equivalent p-value.  

Test your function against this function to ensure you get the same p-value for the scenario at hand (1pt).  

```{r}
prop.test(26, 105, p = 0.15, alternative = "two.sided", correct = FALSE)
# Yes, we get the same p value.
```