---
title: "Lab 1"
author: "Max Sleek"
date: "7/6/2022"
output: 
  html_document:
    number_sections: true
---

Alter the below code chunks to complete the lab. Also, ensure you include your name in the `author` portion of this header. When finished, knit your file to an html and submit this to your TA for grading.  


This first lab is meant to get you acquainted with the data visualization process in R.  We will primarily be using the `Seatbelts` dataset which reports aggregate monthly automobile casualties in Great Britain from the year 1969 to 1979.  

# Initial Inspection 

1) Include the `datasets` package. Then read in the `Seatbelts` dataset, and print the first five rows (0.5pt).  
```{r}
library(datasets)
head(Seatbelts, 5)
```

2) Check the dimension of the data, and find any missing observations (if they exist) (0.5pt).  
```{r}
dim(Seatbelts)
which(Seatbelts == NA)
```

3) Use the `as.data.frame` command to coerce the `Seatbelts` into a dataframe, and store it in a variable called `dat` (0.5pt).  
```{r}
dat <- as.data.frame(Seatbelts)
```

4) Find the median of the `drivers` variable in `dat` (0.5pt).  
```{r}
median(dat$drivers)
```

# Visualization

1) Include the `tidyverse` library, and create a bar plot for the `VanKilled` variable within `dat`.  Add color to the bars of this plot, and label the horizontal and vertical axes "VanKilled" and "Frequency" respectively (2pt).  
```{r}
library(tidyverse)
ggplot(data=dat) +
  geom_bar(aes(x = VanKilled), color = "skyblue", fill = "white") +
  labs(x = "VanKilled", y = "Frequency")
```

2) Next create a scatterplot relating `DriversKilled` to `PetrolPrice` and coloring the points according to `VanKilled`.  Remark as to why it might make sense that the trend appears to be negative (2pt).  
```{r}
ggplot(data=dat) +
  geom_point(aes(x=DriversKilled, y = PetrolPrice, color = VanKilled))
# As the price of oil goes down, gasoline becomes cheaper, meaning more drivers on the road and more possibilities for accidents resulting in death.
```

Next, we will pivot to the `volcano` dataset.  This is a matrix reporting topographical information on the Maunga Whau Volcano.  More specifically, entries report height of the volcano at specific points along the terrain.  I have already included the necessary packages, and I have done some data wrangling (which we will learn about in the coming lectures) to make this data conformable to a heatmap.  

1) *Before* I alter the data via wrangling, check the class, dimension, and type of the volcano matrix (2pt). 
```{r, warning=FALSE}
library(tidyr)
library(tibble)
library(hrbrthemes)
library(dplyr)

#student input
class(volcano)
dim(volcano)
typeof(volcano)
```

2) *After* the data has been manipulated, create a heatmap plotting X against Y and filling according to Z.  Include horizontal and vertical axis labels as well as a title.  In terms of X and Y coordinates, report the approximate position of the crater (spout) of the volcano (2pt).  
```{r, warning=FALSE}
# Heatmap 
volcano %>%
  
  # Data wrangling
  as_tibble() %>%
  rowid_to_column(var="X") %>%
  gather(key="Y", value="Z", -1) %>%
  
  # Change Y to numeric
  mutate(Y=as.numeric(gsub("V","",Y))) %>%
  
  ggplot() %>% +
    geom_tile(aes(X, Y, fill = Z)) +
    labs(x= "X Coordinate", y = "Y Coordinate", title = "Volcano Hieght") +
    scale_fill_distiller(palette = "Spectral")
              
# Find the spout: I approximated the coordinates to be 33,30
volcano[30, 33]
# Guess: 149
```