---
title: "Lab 5"
author: "Max Sleek"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  html_document:
    number_sections: true
---

This lab is meant to give you a bit more practice with writing user-defined functions.  More precisely, this will give you an appreciation for the amount of thought that goes into writing even our most basic functions such as `median`.  You will need to remove the `eval = FALSE` header in some code chunks to get them to execute once you have filled in the student input.  That is, I have given you a template of the overall design of the function.  Repace areas of all caps student input to get the code to run as desired.  

# 

Write a function called `Median` that takes vector input and outputs the median of said quantity.  **Hint: Your function should consider whether the vector is of even length or odd length as we have different ways of handling the median in each case.  I have included a template with some holes.  Your job is to fill in the holes.** (2pt)

```{r, eval = TRUE}

Median <- function(data){
  sorted<-sort(data)
  if (length(data) <= 2){
    Median<-mean(data)
  } else{
    if (length(data) %% 2 == 0) {
      middle_index<-length(data)/2
      middle_two_indices<-c(middle_index, middle_index+1)
      middle_two_numbers<-sorted[middle_two_indices]
      Median<-mean(middle_two_numbers)
    } else{
      index <- ceiling(length(data)/2)
      Median <- sorted[index]
    }
  }
  return(Median)
}

```


#

Test your function on the below sample vectors.  You can use the built-in `median` function as a sanity check.  (1 pt)

```{r}
w <- c(100, 90, 115, 130, 103, 140, 90, 98)
Median(w)
median(w)
x<-c(3,6,5,7,9)
Median(x)
median(x)
y<-c(4, 8, 10, 13, 2, 1)
Median(y)
median(y)
z<-c(10,12)
Median(z)
median(z)
```

#

Now write a function called `skew` that takes in a vector and computes the difference between mean and median.  If said difference is non-zero it should report the corresponding direction of skewness (left or right), otherwise it should report "symmetric".  It should also return the difference itself. (2pt)

```{r, eval = TRUE}

skew <- function(data){
  difference <- mean(data) - median(data)
  if (difference > 0){
    print("Right Skewed")
  }else if (difference < 0 ){
    print("Left Skewed")
  }else{
    print("Symmetric")
  }
  return(difference)
}

```

#

Test your function on the below sample vectors.  (1pt)

```{r}

a<-c(2,3,4,5,6)
skew(a)
b<-c(100, 2, 5, 3, 8)
skew(b)
c<-c(-9, 15, 11, 10)
skew(c)

```

#

Write a function that counts the missing values in a matrix.  You should also have an optional input `encode` that the user can specify if they believe NA's to be encoded as something other than NA.  I will show you how to account for this optional argument; your job is to fill in the logic of the for loops.  (3pts)


```{r, eval = TRUE}

missing <- function(matrix, encode = NULL){
  counter<-0
  if (!is.null(encode)){
    for (i in 1:dim(matrix)[1]) {
      for (j in 1:dim(matrix)[2]){
        if (matrix[i,j] == encode || is.na(matrix[i,j])){
          counter = counter + 1
        }
      }
    } 
  }else{
    for (i in 1:dim(matrix)[1]) {
      for (j in 1:dim(matrix)[2]){
        if (is.na(matrix[i,j])){
          counter = counter + 1
        }
      }
    }
  }
  return(counter)
}


```

#

Test your function on the below two matrices. You should test both specifying the optional argument and not specifying. (1pt)

```{r}

c1<-c(1, NA, 2)
c2<-c(3, 5, 6)
c3<-c(-999, 4, 3)
mat<-cbind(c1,c2,c3)
missing(mat)
missing(mat, encode = 3)

c4<-c(9, -999, NA)
mat2<-cbind(c1,c2,c3,c4)
missing(mat2)
missing(mat2, encode = 9)
```

