---
title: "320 Homework 5"
author: "Max Sleek"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse) 
library(dplyr)
library(tidyr)
```

This homework will familiarize you with the techniques surrounding cleaning data.  More specifically, we will merge and tidy previously disparate and unwieldy data sets.  I have created three such artificial data sets for your consideration.  They list demographic, treatment, and outcome information for up to five patients at a fictitious clinic.  

```{r}
demographics <- data.frame(
  id = c("P1", "P2", "P3"),
  age = c(40, 54, 47),
  country = c("GER", "JPN", "BRA"),
  stringsAsFactors = FALSE
)

treatment_info <- data.frame(
  id = c("P1", "P3", "P4", "P5"),
  control_group = c(0,1,0,1),
  treatment_group = c(1,0,1,0),
  dose = c(0.5, 0, 0.9, 0)
)
adverse_events <- data.frame(
  id = c("P1", "P1", "P3", "P4"),
  ailment = c("Headache", "Neutropenia", "Constipation", "Tachycardia"),
  onset_date = c("2020-12-03", "2021-01-03", "2020-11-29", "2021-01-27"),
  stringsAsFactors = FALSE
)

demographics
treatment_info
adverse_events
```

# Joins 
Rather than viewing these data sets in isolation, we might aim to stitch together the shared information into a holistic conception of patient experience.  

## 

First `left_join` demographics to treatment_info by the shared key.  Store this new dataframe as `demo_treat`.  

### 

Display the entirety of `demo_treat`. 

```{r}
demo_treat <- left_join(treatment_info, demographics, by = "id")
demo_treat
```

### 
Now repeat the same process, but `inner_join` the two original data sets rather than `left_joining` them.  Call this new data set `DEMO_TREAT`. Print the entirity of this dataframe as well. 

```{r}
DEMO_TREAT <- inner_join(treatment_info, demographics, by = "id")
DEMO_TREAT
```

### 

Comment on the number of rows in `demo_treat` and `DEMO_TREAT`.  Why are they not equal?  

The number of rows is different because inner join does not allow for missing values; since it is sorted by id, only P1 and P3 are present since those are the only indexes present in both data sets. For the left_join function, the values of demographics that match the indexes in treatment info are essentially being added to the treatment info dataset. 

## 

**inner_join** `adverse events` to `DEMO_TREAT`.  Store this as `Final`.  Would we expect a different result if we **inner_joined** to `demo_treat`? 

```{r}
Final <- inner_join(adverse_events, DEMO_TREAT, by = "id")
```

We would expect a different result if we inner_joined to demo_treat because P4 would be considered a matching index and therefore would be included in the join funtion. 

# Tidy

## 

`Final` is not a tidy data set because one of the original datasets was also untidy.  Which original data set was untidy and explain why this was the case.  

The treatment_info data set could be considered untidy because the control group and treatment group are in different columns. We can merge them into one column titled `group`.

## 

Correct `Final`, and call this `Tidy_Final.  Print the tidy data set to the screen.  *Hint: If you have identified the correct data frame and issue, the fix doesn't even require a tidy command.  Rather, it should just collapse information into a preexisting column and delete extraneous ones.*

```{r}
for (i in 1:(count(Final))){
  if (Final$control_group[i] == 0){
    Final$group[i] <- "Treatment"
  }
  else{
    Final$group[i] <- "Control"
  }
}
Final <- Final %>% subset(select = -c(4,5))
Final <- relocate(Final, group, .before = dose)
```

##

A more direct application of the tidy commands can be seen when applying them to `table2` below.  Take particular care to observe the `type` variable.  

```{r}
head(table2)
```

Tidy up `table2` using one of the tidy commands.  

```{r}
table2 <- table2 %>% pivot_wider(names_from = type, values_from = count)
```